---
title: "a1751699_A3"
author: "Zhao Ming Soh"
date: "19/05/2022"
output:
  html_document:
    df_print: paged
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
pacman::p_load(tidyverse, gglm, broom)
```

## a) Read the data in
```{r}
lung_cancer <- read_csv("lung_cancer.csv")
lung_cancer
```

## b) Perform EDA
```{r}
skimr::skim(lung_cancer)

lung_cancer %>%
  ggplot(aes(cases)) + geom_bar()

lung_cancer %>%
  ggplot(aes(pop, cases)) + geom_point()

lung_cancer <- lung_cancer %>%
  mutate(
    across(where(is.character), factor)
    )

# Boxplot that shows the relationship between cases in the age and city groups
lung_cancer %>%
  select(where(is.factor),cases) %>%
  pivot_longer(-cases) %>%
  ggplot(aes(value,cases)) +
  geom_boxplot() + 
  facet_wrap(~name, scales = "free_x")

# Boxplot that shows the relationship between proportion of cases to populations in the age and city groups
lung_cancer %>%
  mutate(
    p = cases/pop
  ) %>%
  select(p,where(is.factor)) %>%
  pivot_longer(-p) %>%
  ggplot(aes(value,p)) +
  geom_boxplot() + 
  facet_wrap(~name, scales = "free_x")
```

## c) Fit a Poisson Rate Regression (M1)
```{r}
M1 <- glm(cases ~ 1, family = poisson, offset = log(pop), data = lung_cancer)
summary(M1)
```

## d) Fit a Poisson Rate Regression (M2)
```{r}
M2 <- glm(cases ~ age + city, family = poisson, offset = log(pop), data = lung_cancer)
summary(M2)
```

## e) Fit a Poisson Rate Regression (M3)
```{r}
M3 <- glm(cases ~ age + city + log(pop), family = poisson, offset = log(pop), data = lung_cancer)
summary(M3)
```

## f) Use Anova to compare M1 and M2
```{r}
anova(M1,M2, test = "LRT")
anova(M2, test = "LRT")
```
There is a significant difference between M1 and M2 as shown by the anova test where M2 has a p-value < 0.05. Therefore, both city and age are significant predictors in predicting the number of cases of lung cancer.

## g) Find AIC of all 3 models
```{r}
AIC(M1,M2,M3)
AIC(M2) - AIC(M3)
```
The difference in AIC between M2 and M3 is 1.95 which is within 2 of each other and the degree of freedom of M2 is lower than M3. Therefore, the rule of thumb is to choose the smaller model when this happens.

## h) Summary of coefficients for M2
```{r}
coef_M2_age55_59 <- 1.1010
exp(1.1010)

```
The coefficient of the age group 55-59 is 1.1010, this is equivalent to an increase of 3.007172 times (after taking the exponential on the coefficient e^1.1010 = 3.007172) for each increase in the population of the age group 55-59 by 1 person as compared to the age group 40-54 keeping everything else constant.

## i) Obtain the Pearson Residuals for M2
```{r}
lung_cancer <- lung_cancer %>%
  add_column(
    M2_res = residuals(M2, type="pearson"),
    M2_fit = fitted(M2)
  )
```

## j) Plot residuals vs
```{r}
lung_cancer %>%
  select(where(is.double)) %>%
  pivot_longer(-M2_res) %>%
  ggplot(aes(value, M2_res)) + geom_point() + facet_wrap(~name, scale = "free")

lung_cancer %>%
  select(where(is.factor), M2_res) %>%
  pivot_longer(-M2_res) %>%
  ggplot(aes(value, M2_res)) + geom_boxplot() + facet_wrap(~name, scale = "free")

```

## k) 
```{r}
new_data<-tibble(age = "40-54", pop = 4000, city = "Fredericia")
lambda<-predict(M2,newdata = new_data, type = "response")
ppois(5, lambda)

# Trial and Error with different Lambda values
ppois(5, 10)
ppois(5, 5)
```

The log additive (multiplicative) model

$$\begin{aligned}
\mu_{ijk} = n_{ijk}e^{\alpha_i+\beta_j+\gamma_k}
\end{aligned}$$

